---
title: "BDTT"
author: "Florent Mazel (helped by a friendly reviewer)"
date: "February 14, 2017"
output:  html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```
  

**In this script, we provide:**
  
1. A general description of the analysis
2. Examples of the use of BDTT on a dataset  
 2.1. A rapid analysis, using the dataset provided in the picante package  
 2.2. A more detailed analysis, using the same dataset provided in the picante package 

#General workflow of the analysis

**Input DATA**  

  - site X species matrix (or 'samples/hosts X unique 16S sequences' in microbiology -- The 100% similarity OTU table)  
  - the species phylogeny (the names of the leaves must match those in the site X species matrix)  
  - site * sites environmental distances (e.g. climatic distances, dietary or phylogenetic   distances between hosts, etc). 

**Computational steps (n is the total number of slices defined)**

1. Create n Branch * Sites Matrices (with function 'GetBranchOcc') (or n OTU tables)
2. Compute n beta-diversity matrices (function 'GetBeta')
3. Compute the Mantel correlation profile along the phylogenetic scale, i.e. along the n slices (function 'GetCorrelations')

**Output DATA**   

  - n Branch * Sites Matrices (or n OTU tables)  
  - n beta-diversity matrices   
  - a correlation profile along the phylogenetic scale  

**NOTE 1** .  Results are successively saved in the specified output directory, and are not created in the R env.  (to avoid memory problems when a lot of slices are defined and/or when big site * species matrices are used)  

**NOTE 2** . Instead of using 'lapply' as in the following examples, one can use 'mclapply' to parrallelize the computations  

# Examples to test the analysis


**Load** the data, packages and functions

```{r,warning=F,message=F}
# Libraries
library(picante) #loaded to use the example dataset
library(ape) # to get the Branch * sites matrices
library(abind)
library(caper) # to get the Branch * sites matrices
library(betapart) # to compute Beta-Diversity matrices
library(ecodist) # to compute correlations profiles (function 'MRM')
library(bigmemory)
library(knitr)

source("BDTT_functions.R") #load all associated functions, used to run BDTT

# Examples
data(phylocom)
TreeExmple=phylocom$phylo
plot(TreeExmple)

SiteSpExmple=phylocom$sample
SiteSpExmple

Env=rnorm(n=dim(SiteSpExmple)[1],mean=0) #random environment
names(Env)=rownames(SiteSpExmple)
ENvdi=as.matrix(dist(Env))
```

#Rapid test            

**NOTE:** In the following, replace "" with your own path to your appropriate directory.

```{r}
slices=seq(from=0,to=5, by=0.5) # first, time slices are defined

a<-lapply(slices,GetBranchOcc,tree=TreeExmple,sitesp=SiteSpExmple,pathtoSaveBranchOcc="",bigmatrix=F)
a<-lapply(slices,GetBetaDiv,pathtoGetBranchOcc="",pathtoSaveBeta="")
Cors=sapply(slices,GetCorrelations,indice="sor",EnvDist=ENvdi,pathtoGetBeta="",TypeofMantel="Spearman",nperm=1000)
colnames(Cors)=slices
```

**Finally, we can use a simple plot to represent the results:**
  
```{r}
plot(slices,Cors[1,],type='b',xlab='Phylogenetic scale',ylab='R2')
```

#Detailed test           

####A. Get Branch*Sites Matrices

**Function** 'GetBranchOcc'  

  - computes a Branch*Sites matrix and directly save it in a file (not in Renv)  

**Input variables**  

  -  slice: the age of the desired slice
  -  tree: the community phylogenetic tree (must be ultrametric)
  -  sitesp: the site * species matrix, column names must match to the leaf names of the phylogenetic tree (In microbiology, it represents the OTU table with 100% similarity OTUs)
  -  pathtoSaveBranchOcc: directory where the Branch*sites matrices are saved
  -  bigmatrix=F: if site*species is a VERY big matrix, it is highly recommended to set bigmatrix=T (use of the bigmemory package)

**Output** 

 - save a branch*site matrix (or OTU table) in 'pathtoSaveBranchOcc'

For a given slice  
```{r}

GetBranchOcc(slice=1,tree=TreeExmple,sitesp=SiteSpExmple,pathtoSaveBranchOcc="",bigmatrix=F)
# To get the full vector of slices
a=lapply(seq(from=0,to=5, by=0.5),GetBranchOcc,tree=TreeExmple,sitesp=SiteSpExmple,pathtoSaveBranchOcc="",bigmatrix=F)

```

####B. Get Beta Diversity matrices along all slices

**Function**  'GetBetaDiv'

 - computes Beta-diversities from a site*species matrix and directly saves the matrix of beta-diversities

**Input variables**  

  -  slice: the age of the desired slice  
  -  pathtoGetBranchOcc: the input file where the branch*sites matrix is saved (result of the   -  function GetBranchOcc)  
  -  pathtoSaveBeta: the output file where the matrices of beta diversities are saved (several beta-diversity metrics are used)  
  -  bigmatrix=F: if you used bigmatrix=T to create the branch*sites matrix with 'GetBranchOcc' (OTU table), use bigmatrix=T again.  

**Output** 

  -  save Beta diversity matrices as a 3D array in 'pathtoSaveBeta' 
  -  Array = array of sites*sites*beta diversity metrics 
    -  all beta diversity metrics:  
    -  "jtu": True Turnover component of Jaccard (Presence/Absence)  
    -  "jne": Nestedness component of Jaccard (Presence/Absence)  
    -  "jac": Jaccard (Presence/Absence)  
    -  "sne": Nestedness component of Sorensen (Presence/Absence)  
    -  "sor": Sorensen (Presence/Absence)  
    -  "bctu": True Turnover component of Bray-Curtis (Abundance version of Sorensen)  
    -  "bcne": Nestedness component of Bray-Curtis (Abundance version of Sorensen)  
    -  "bc": Bray-Curtis (Abundance version of Sorensen)  


**Run for one given slice** 

```{r}
GetBetaDiv(slice=1,pathtoGetBranchOcc="",pathtoSaveBeta="")
```

**Run for multiple given slice** 
```{r}
slices=seq(from=0,to=5, by=0.5)
a=lapply(slices,GetBetaDiv,pathtoGetBranchOcc="",pathtoSaveBeta="")
```

**Example**
```{r}
load("BetaDiv_BetaDivSliceNo5.rdata")
dimnames(Betaa)
```

#### C. Get the correlation profile

**Function** 'GetCorrelations'

  -  computes correlations between beta-diversities and environmental distances at the desired slice

**Input variables**  

  -  slice: the age of the desired slice  
  -  indice: the betadiv metric you want  
  -  pathtoGetBeta : the file where Beta-Diversity matrices were previously stored (output of the 'GetBetaDiv' function)   
  -  EnvDist: matrix of environmental distances. (e.g. geographic or climatic distances between sites, dietary or phylogenetic distances between hosts, etc).  
  -  TypeofMantel: type of correlation used to run the Mantel test (either "Spearman" or "Pearson")   -  nperm: number of permutations to perform to compute a p-value  

**Output** 

  -  A 2*n matrix with R2 coefficients and their associated p-values


**Run for one given slice** 

```{r}
GetCorrelations(slice=1,indice="sor",EnvDist=ENvdi,pathtoGetBeta="",TypeofMantel="Spearman",nperm=1000)
``` 

**Run for multiple given slice** 

```{r}
slices=seq(from=0,to=5, by=0.5)
Cors=sapply(slices,GetCorrelations,indice="sor",EnvDist=ENvdi,pathtoGetBeta="",TypeofMantel="Spearman",nperm=1000)
colnames(Cors)=seq(from=0,to=5, by=0.5)
Cors
``` 

**Finally, we can use a simple plot to represent the results:**

```{r}
plot(slices,Cors[1,],type='b',xlab='Phylogenetic scale',ylab='R2')
```

```{r,eval=F}

