## Table of contents

#### 1. [The Packages](#Packages)

#### 2. [The BDTT function](#Functions)

#### 3. [Example](#Example)

#####3.1. Computing BDTT

#####3.2. Testing the statistical link between BDTT and metadata



## 1. The needed R packages <a name="Packages"></a>

Before running the BDTT analysis, you will need to install and load the following R packages: 

* ape
* castor
* matrix 
* abind

```{r, message=FALSE}
library(ape)
#library(betapart)
library(castor)
library(abind)
library(Matrix)
```


## 2. The BDTT function <a name="Functions"></a>

##### The BDTT function 
requires the follwing inputs: 

###### similarity_slices:
the slices (i.e. the multiple phylogenetic resolutions) at which you want to aggregate the tips of the phylogeny and compute corresponding beta-diversity. 0 corresponds to no aggregation, i.e use the raw tips of the phylogeny as microbial units. Values >0 will aggregate the tips of the phylogeny according to the given value to create aggregated microbial units and compute corresponding beta-diversity. Use the 'getHnodes' function to have an idea of the resolution slices you can explore (see below). 

###### tree: 
the species (or OTUs, or sequence variants) phylogeny (the names of the tips must match those in the site*species matrix)

###### sampleOTUs:
samples * species (or OTUs, or sequence variants) matrix

###### onlyBeta: 
Putting "TRUE" (default) will make the function return beta-diversity dissimilary matrices only
Putting "FALSE" will make the function return beta-diversity dissimilary matrices + matrix detailling the relationship between tips and the aggregated units. 

###### metric:
Beta-diversity metric chosen; we provide Jaccard ("Jac") its true turnover component ("Jac_TT") and Bray-Curtis ("Bray"). 

The function requires the following input: 

###### tree: 
the species (or OTUs, or sequence variants) phylogeny

## 2. Examples <a name="Example"></a>

### Computing BDTT

```{r, message=FALSE}
library(picante)

data(phylocom)
TreeExample=phylocom$phylo
plot(TreeExample)

SiteSpExample=t(phylocom$sample)
SiteSpExample

source("BDTT_functions.R")

hist(get_all_node_depths(TreeExample))
slices=c(0:3)
Betas=BDTT(similarity_slices = slices,tree = TreeExample,sampleOTUs = (SiteSpExample))
```


### Linking BDTT with environement / metadata

Create random metada catageory 

```{r, message=FALSE}
Meta=sample(x=c("Condition_1","Condition_2"),size=dim(SiteSpExample)[2],replace = T) 
names(Meta)=colnames(SiteSpExample)
Meta
```

Test statistically the link between metadata and BDTT using PERMANOVA

Load vegan to be able to use adonius function

```{r, message=FALSE}
library(vegan)
```

Example of the test for a given resolution (0) and a given metric (Jaccard); make sure that samples are in the same order

```{r, message=FALSE}
samples=names(Meta)
adonis(Betas["0","Jac",samples,samples]~Meta[samples])
```

Construct table to store results in a ready-to-use format for ggplot


```{r, message=FALSE}
predictors="Conditions1_2"
StatsRes=expand.grid(similarity_slices=as.character(slices),predictors=predictors,metric=c("Jac","Bray"))
StatsRes[["F.Model"]]=StatsRes[["R2"]]=StatsRes[["Pr(>F)"]]=NA
head(StatsRes)
```

Run multiple PERMANOVA across phylogenetic resolution and store results in a table ready to use for ggplot

```{r, message=FALSE}
for (i in as.character(slices))
{
   res=unlist(adonis(formula =Betas[i,"Jac",samples,samples]~Meta[samples])$aov.tab[1,c(6,5,4)])
   StatsRes[(StatsRes$metric=="Jac")&(StatsRes$similarity_slices==i),4:6]=res
   res=unlist(adonis(formula =Betas[i,"Bray",samples,samples]~Meta[samples])$aov.tab[1,c(6,5,4)])
   StatsRes[(StatsRes$metric=="Bray")&(StatsRes$similarity_slices==i),4:6]=res
}
```

We can then plot the profiles of R2 along the phylogenetic time scale:

```{r, message=FALSE}
library(ggplot2)
ggplot(aes(y=R2,x=similarity_slices,colour=predictors,group=factor(predictors)),data=StatsRes)+geom_point()+geom_line()+facet_wrap(~metric)
```

or just the profile for the significant effects

```{r, message=FALSE}
ggplot(aes(y=R2,x=similarity_slices,colour=predictors,group=factor(predictors)),data=StatsRes[StatsRes$`Pr(>F)`<.05,])+geom_point()+geom_line()+facet_wrap(~metric)
```

